---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Sebastien Leblanc
## Loading and preprocessing the data

### Loading the data & necessary libraries
```{r}
library(lubridate)
library(dplyr)
library(ggplot2)

act_data = read.csv("activity.csv")
```
## What is mean total number of steps taken per day?

### Histogram of total number of steps (NA data ignored)
```{r}
daily_avrg <- act_data %>%
    group_by(date) %>%
    select(steps) %>%
    summarize(steps_sum = sum(steps, na.rm=T))
ggplot(daily_avrg, aes(steps_sum))+
    geom_histogram(binwidth=1000)
```

### Mean and median of total number of steps 
```{r}
mean(daily_avrg$steps_sum)
median(daily_avrg$steps_sum)
```

## What is the average daily activity pattern?

### Time series plot of daily activity (NA values ignored)
```{r}
interval_avrg <- act_data %>%
    group_by(interval) %>%
    select(steps) %>%
    summarize(steps_mean = mean(steps, na.rm=T))
ggplot(interval_avrg, aes(interval,steps_mean))+
    geom_line()
```

### Identifying the interval with the most steps

```{r}
interval <- which.max(interval_avrg$steps_mean)
max_steps <- max(interval_avrg$steps_mean)
```
The interval showing the highest number of steps is `r interval` with a maximum average of `r max_steps` steps.


## Imputing missing values

### Total number of missing values
```{r}
total_nas <- sum(is.na(act_data))
ratio_nas <- sum(is.na(act_data))/nrow(act_data)*100
```

A total of `r total_nas` values are missing from the dataset, representing `r ratio_nas`% of the data.

### Filling in the missing values
The missing values are replaced by the average of the corresponding time interval.
```{r}
impute.mean <- function(x){
    replace(x, is.na(x), mean(x, na.rm = TRUE))
}
act_imputed <- act_data %>%
    group_by(interval) %>%
    mutate(steps=impute.mean(steps)) %>%
    ungroup() 
```

### Histogram of the daily total steps
The histogram shows the density distribution of the total number of steps per day.
```{r}
act_imp_daily  <- act_imputed %>%
    group_by(date) %>%
    select(steps) %>%
    summarize(steps_sum = sum(steps))
ggplot(act_imp_daily, aes(steps_sum))+
    geom_histogram(binwidth=1000)
```

#### Compute the mean and median of total daily steps
```{r}
mean(act_imp_daily$steps_sum)
median(act_imp_daily$steps_sum)
```

## Are there differences in activity patterns between weekdays and weekends?

### Factoring weekdays and weekends
```{r}
act_imputed$date <- ymd(act_imputed$date)
act_imputed <- act_imputed %>%
    ungroup() %>%
    mutate(
        week=factor(
            ifelse(wday(date) == 0 | wday(date) == 6,
                   'weekend','weekday')
        )
    ) %>%
    group_by(week, interval) %>%
    select(steps) %>%
    summarize(steps_mean = mean(steps))
```

### Panel plot showing weekdays and weekend patterns
```{r}
ggplot(act_imputed, aes(interval, steps_mean))+
    geom_line()+
    facet_grid(week ~ .)
```
